---
title: "Avatar Data Set Analysis"
author: "Avatar Fan"
date: '2020-12-08'
output:
  html_document:
    df_print: paged
  pdf_document: default
categories: []
tags: []
slug: avatar-data-set-analysis
---

Avatar the Last *Airbender* is my most favorite show of all time, with a world that feels alive,
characters that are fun, and magic that feels natural and fun. I am going to use the
avatar.csv data set, which was found on Tidy Tuesday. The data consist of 11 variables and
13,385 observations. 
The **id** variable is an integer that is purely used to uniquely identifyeach row. 
The **book** is a character that shows which book, or season, the line is in. 
The **book_num** is the book number, showing whether the line is in season 1, 2, or 3. 
The **chapter** is the name of the episode. 
The **chapter_num** is the episode number in relation to the series. 
The **character** is the character speaking or acting during the scene, or whether it is a scene description. 
The **full_text** is the full text of the scene, ie. the scene description and character text. 
The **character_words** is the text coming from the characters if they speak, if they do not, then it is NA. 
The **writer** the writer, or writers, of the book/episode.
The **director** the director of the episode. 
The **imbd_rating** the IMDB rating for the episode, or how people thought about it.

```{r, message=FALSE,results = FALSE}
library(tidyverse)
tuesdata <- tidytuesdayR::tt_load('2020-08-11')
tuesdata <- tidytuesdayR::tt_load(2020, week = 33)
avatar <- tuesdata$avatar
```

## "Special Move!"

In other kids shows, like *Power Rangers*, the main characters and will yell say their special moves often, enough to have it ingrained into your head. How often does *Avatar: The Last Airbender* do this, and who does it the most?

```{r, message=FALSE,results = FALSE}
library(tidyverse)
library(tvthemes)

Katara <- avatar %>%
  filter(!is.na(character_words)) %>%
  filter(character == "Katara") %>%
  filter(str_detect(character_words, regex("waterbend", ignore_case = TRUE))) %>%
  group_by(character) %>%
  summarise(special = n())
Zuko <- avatar %>%
  filter(!is.na(character_words)) %>%
  filter(character == "Zuko") %>%
  filter(str_detect(character_words, regex("firebend", ignore_case = TRUE))) %>%
  group_by(character) %>%
  summarise(special = n())
Aang <- avatar %>%
  filter(!is.na(character_words)) %>%
  filter(character == "Aang") %>%
  filter(str_detect(character_words, regex("airbend", ignore_case = TRUE))) %>%
  group_by(character) %>%
  summarise(special = n())
Toph <- avatar %>%
  filter(!is.na(character_words)) %>%
  filter(character == "Toph") %>%
  filter(str_detect(character_words, regex("earthbend", ignore_case = TRUE))) %>%
  group_by(character) %>%
  summarise(special = n())
Sokka <- avatar %>%
  filter(!is.na(character_words)) %>%
  filter(character == "Sokka") %>%
  filter(str_detect(character_words, regex("boomerang", ignore_case = TRUE))) %>%
  group_by(character) %>%
  summarise(special = n())
dat <- rbind(Sokka, Aang, Katara, Zuko, Toph)

ggplot(data = dat, aes(x = character, y = special, fill = character)) +
  geom_bar(aes(reorder(character, special)), stat = "identity") +
  scale_fill_avatar("AirNomads")
```
We took the data and filtered it specifically to only when they say their specific power. For Katara we did waterbending, Aang had airbending, Zuko had firebending, Toph had earthbending, and Sokka had his trusty boomerang. We found that Katara, over the series, says hers the most, with Sokka saying his weapon of choice the least. This may be because of the episodes they were featured in. Katara had multiple episodes where she met waterbending masters and was trained by them, so those episodes were great contributors to her word count. While Sokka used a boomerang at the beginning of the series, later on, he switched to a sword after a training episode, so that might explain why he is at the bottom of the list despite the boomerang being what he is most known for.


## Highest and Lowest Rated Episode for Each Season

*Avatar: The Last Airbender* is a great TV show, but as with anything, there will always be favorites and least favorites. What episodes are the highest and lowest rated, and what is it for each season?

```{r, message=FALSE,results = FALSE}
library(tidyverse)
library(dplyr)

episode <- avatar %>%
  select(imdb_rating,chapter,book) %>%
  distinct() %>%
  filter(!is.na(imdb_rating))

book1 <- episode %>%
  filter(book == "Water") %>%
  arrange(imdb_rating)

book2 <- episode %>%
  filter(book == "Fire") %>%
  arrange(imdb_rating)
book3 <- episode %>%
  filter(book == "Earth") %>%
  arrange(imdb_rating)
bestworst <- rbind(head(book1, 1), tail(book1, 1), head(book2, 1), tail(book2, 1), head(book3, 1), tail(book3, 1))

ggplot(data = bestworst, aes(x = chapter, y = imdb_rating, fill = book)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_avatar("WaterTribe") +
  ggtitle("Best and Worst Episodes Each Season") +
  xlab("Episode Name") +
  ylab("Rating")
```
From the graph, we can see the highest and lowest rated shows for each book, or season. There is a trend in the highest rated episodes that is not very apparent unless you are a fan of the show. "THe Siege of the North", "The Crossroads of Destiny", and "Sozin's Comet" are all the final episode of their respective season. That would explain why those episodes are the highest for their season; they are the climactic end. On the other side, the lowest rated episodes for their perspective season all also have a trend. "Avatar Day", "Nightmares and Daydreams", and "The Great Divide" are all filler episodes, or episodes where the plot moves the slowest. In all these episodes, the characters are not driving the plot forward by having them learn new elements or come up with plans to defeat the Firelord. The characters are instead dealing with "sidequest". For example, "The Great Divide" is the main cast of characters escorting two tribes across a ravine, and in "Avatar Day" the main cast is solving a murder mystery. So these characters are the lowest rated for their seasons because they carry less plot importance than the other episodes.


## Ratings Over Time

Does *Avatar: The Last Airbender* improve or get worst with time? As with any shows, it'll have valleys and peaks, so where do they occur?

```{r, message=FALSE,results = FALSE}
library(patchwork)
library(tvthemes)

seasonrating <- avatar %>%
  select(book_num, chapter_num, imdb_rating) %>%
  distinct()

waterrating <- seasonrating %>%
  filter(book_num == 1) %>%
  filter(!is.na(imdb_rating)) %>%
  ggplot(aes(x = chapter_num, y = imdb_rating)) +
  geom_point() +
  geom_line() +
  xlab("Episode Number") +
  ylab("Rating") +
  ggtitle("Season 1 Ratings")

earthrating <- seasonrating %>%
  filter(book_num == 2) %>%
  ggplot(aes(x = chapter_num, y = imdb_rating)) +
  geom_point() +
  geom_line() +
  xlab("Episode Number") +
  ylab("Rating") +
  ggtitle("Season 2 Ratings")

firerating <- seasonrating %>%
  filter(book_num == 1) %>%
  filter(!is.na(imdb_rating)) %>%
  ggplot(aes(x = chapter_num, y = imdb_rating)) +
  geom_point() +
  geom_line() +
  xlab("Episode Number") +
  ylab("Rating") +
  ggtitle("Season 3 Ratings")

waterrating / earthrating / firerating

```
We can see from the graphs for the episodes and the best and worst graph from the previous question that most seasons follow a trend. At the beginning of the season, the rating is mid, it fluctuates and takes a large dip towards the middle, usually in one of the aforementioned worst episodes, and the climbs to the climax at the final of the season.


## Most Chattiest Characters

Who talks the most in the show? There are many candidates, like Azula and her silver tongue, Katara saying the opening each episode, and Sokka with his witty quips. But who says the most lines in the show?


```{r, message=FALSE,results = FALSE}
library(ggplot2)
library(patchwork)
library(tvthemes)

lines <- avatar %>%
  filter(!is.na(character_words)) %>%
  group_by(character,book) %>%
  summarise(Lines = n()) %>% 
  arrange(Lines, descending = TRUE)

waterlines <- lines %>%
  filter(book == "Water") %>%
  tail(n = 5) %>%
  ggplot(aes(x = character, y = Lines, fill = character)) +
  geom_bar(aes(reorder(character, Lines)), stat = 'identity') +
  theme_avatar() +
  coord_flip() +
  scale_fill_avatar(palette = "WaterTribe") +
  theme(legend.position = "none") +
  xlab("# of Lines") +
  ylab("Character") +
  ggtitle("Chattiest Character: Season 2")

earthlines <- lines %>%
  filter(book == "Earth") %>%
  tail(n = 5) %>%
  ggplot(aes(x = character, y = Lines, fill = character)) +
  geom_bar(aes(reorder(character, Lines)), stat = 'identity') +
  theme_avatar() +
  coord_flip() +
  scale_fill_avatar(palette = "EarthKingdom") +
  theme(legend.position = "none") +
  xlab("# of Lines") +
  ylab("Character") +
  ggtitle("Chattiest Character: Season 2")

firelines <- lines %>%
  filter(book == "Fire") %>%
  tail(n = 5) %>%
  ggplot(aes(x = character, y = Lines, fill = character)) +
  geom_bar(aes(reorder(character, Lines)), stat = 'identity') +
  theme_avatar() +
  coord_flip() +
  scale_fill_avatar(palette = "FireNation") +
  theme(legend.position = "none") +
  xlab("# of Lines") +
  ylab("Character") +
  ggtitle("Chattiest Character: Season 3")

waterlines / earthlines / firelines
```

At the beginning of the show, we can see that Aang, Katara, and Sokka did most of the talking, as they were the main characters and the ones talking the most. Zuko and Iroh were also there, but they were mostly side characters or antagonist for the most part. We can see in the second season that the three main characters talk the most, but this time Sokka takes the number one spot. Toph is also introduced this season so she is also on the top 5, and weirdly enough, Iroh is on the list and not Zuko. In the third season, Sokka still takes the top spot, but Katara has been moved down and Zuko takes top 3 thanks to him joining the main team. Iroh has been kicked off for Toph, but that makes sense, since everyone in the top 5 are part of the main team.


## Best Writer(s) Award (According to the Viewers)

The writers of this show are some of the best. They are able to write a compelling story line while adding funny jokes and character growth along the way. But who, or which group, has the highest average rated episodes?

```{r, message=FALSE}

writer <- avatar %>%
  select(book_num, chapter_num,writer, imdb_rating) %>%
  group_by(writer) %>%
  summarise(score = mean(imdb_rating)) %>%
  filter(!is.na(score))  %>%
  arrange(score)
writer
ggplot(data = tail(writer, 10), aes(x = writer, y = score, fill = writer)) +
  geom_bar(aes(reorder(writer, score)), stat = 'identity') +
  xlab("Average IMDB Score") +
  ylab("Writers") +
  scale_fill_discrete(labels = c("A. Ehasz, DiMartino, Konietzko","Group of Writers", "Konietzko, DiMartino", "DiMartino, Konietzko", "E. Ehasz, Hedrick", "DiMartino", "E.Ehasz", "Hedrick", "Chan", "E.Ehasz, Hedrick, A.Ehasz, O'Bryan")) +
  ggtitle("Average IMDB Score for Each Writer(s)") +
  theme(axis.text.x = element_blank())
```
We can see that the best episodes are written by a group of writers. Interestingly, We can see two people with the same last name, Ehasz. They are actually married and helped write the show together. Interestingly, one of the highest rated episodes created by a group of writers is an episode called "Tales from Ba Sing Se". It is an episode that is a collection of short stories, which explains the high volume of writers for that particular episode, and the high ratings the episode received.

##Conclusion
The Avatar: The Last Airbender data set has so much to look into thanks to its extensive data. We found out that despite the show being called "The Last Airbender", airbending is not mentioned as much from the airbender as waterbending is from the waterbender. From studying episodes, we can find the general trend that each season takes, where it starts out mid, takes a dip in the middle due to a filler episode, and reaches a high point in ratings during the climactic final. We can see that the most chattiest characters Aang and Sokka, two characters that are prevalent in all three seasons. And from looking at the ratings from writers, we can see that while there are standout writers, the best episodes are from a combination of writers.


