---
title: "Password Data Analysis"
author: "William A. Bounsavy"
date: '2020-12-08'
slug: password-data-analysis
categories: []
tags: []
---

The password data set from tidytuesday contains a list of many commonly used passwords.

```{r}
library(tidyverse)
tuesdata <- tidytuesdayR::tt_load('2020-01-14') 
tuesdata <- tidytuesdayR::tt_load(2020, week = 3)
passwords <- tuesdata$passwords
passwords 
```


## Easiest Passwords to Crack

The data set may sort the data set on most common, but what about easiest to crack. Probably the first 10, since they are the most common, but what about the rest?

```{r}
crack <- passwords %>%
  tail(497) %>%
  filter(time_unit == "minutes") %>%
  arrange(value) %>%
  mutate(seconds = value*60)

seconds <- passwords %>%
  tail(497) %>%
  filter(time_unit == "seconds") %>%
  arrange(value)

seconds
crack
```

We can see with the two data sets that a lot of the passwords that are easily crackable have around the same value to be cracked.

## Easiest Passwords to Crack

The passwords data set also gives each password a category, as in whether it is password-related, an animal, or simple-alphanumeric. What category is the most commonly used?

```{r}
dat <- passwords %>%
  filter(!is.na(category)) %>%
  group_by(category) %>%
  summarise(count = n())

ggplot(data = dat, aes(x =  category, y = count)) +
  geom_bar(aes(reorder(category, count)),stat = "identity") +
  coord_flip()
```
The category commonly used and is easiest to crack seems to be the name category, it almost doubles the cool-macho category. This makes sense, peoplelike to use names when they are making their password. Another interesting thing is that simple-alphanumeric passwords are also at the top of the list. This is probably due to them being easy to remember, things like 1111 and 2222 and 1234 can be easily written down and remembered. Phones use those passwords a lot, in fact.

## Strongest Password

We know which categories are the most common and the easiest ones to crack, but what about the strongest password? Each password in the data set comes with their own separate strength score, so which one is the strongest?

```{r}
strength <- passwords %>%
  arrange(desc(strength)) %>% 
  head(15)

weak <- passwords %>%
  arrange(desc(strength)) %>% 
  tail(492) %>%
  head(25)

ggplot(strength, aes(x = password, y = strength)) +
  geom_bar(aes(reorder(password, strength)), stat = "identity") +
  coord_flip()

ggplot(weak, aes(x = password, y = offline_crack_sec)) +
  geom_bar(aes(reorder(password, strength)), stat = "identity") +
  coord_flip()

strength
weak
```
We can see that in the top 15 strongest passwords on the list there is a trend. For the first few, the strongest passwords are simple passwords with just a word and then numbers after it. This makes sure that the password is easy to remember but hard to break.

#Conclusion

TBD